
## R code for analysis of the Arroyo Hondo Pueblo turkey isotope and faunal data.
 
Cyler Conrad, Department of Anthropology, University of New Mexico, cylerc@unm.edu

Ben Marwick, Department of Anthropology, University of Washington, bmarwick@uw.edu

This document contain R code to reproduce the plots and statistical analysis presented in 

> Conrad, C., Jones, E.L., Newsome, S.D. and Schwartz, D.W. (in review).  Eggshell, Bone Isotopes and Turkey Husbandry at Arroyo Hondo Pueblo.  Journal of Archaeoloical Science: Reports.

All data required to perform the analyses can be found at XXXX (Conrad et al. XXXX). The development version of this document can be found at https://github.com/cylerc/JASR_AH

Details of the computational environment and software dependencies are listed at the end of this document. 
 
```{r setup}
# set the base directory for knitr to the directory above this one
library(knitr)
opts_knit$set(root.dir = '../')
```

```{r load_libraries, message = FALSE, warning = FALSE}
# see the output of sessionInfo() at the bottom for package version numbers
library(ggplot2) 
library(reshape2)
library(vegan)
library(knitr)
library(plyr)
library(dplyr)
```

```{r relative_abundance}
# Calculate relative abundance. First, upload the Arroyo Hondo NISP table.
# Second, take all turkey NISP values and divide by the column totals to calculate
# the relative abundance of turtles per stratum. 

ah <- read.csv("data/ahNISP.csv", stringsAsFactors = FALSE, check.names = FALSE)
kable(ah)

turkey_NISP <- ah[ah$Taxon == "Meleagris gallopavo", -1]
column_totals <- colSums(ah[ ,-1])
relative_abundance <-  turkey_NISP / column_totals
relative_abundance <- t(relative_abundance)
relative_abundance <- data_frame(`Date (C.E.)` = row.names(relative_abundance), 
                                 `Relative Abundance` = relative_abundance[,1])

# save as CSV 
write.csv(relative_abundance, "data/fig6a.csv", row.names = FALSE)

# display table
kable(relative_abundance)
```

```{r plot_relative_abundance}
# Plot of relative abundance. Use relative abundance
# values that were previously calculated to create plot of changing abundance 
# through time. 

ra <- read.csv("data/fig6a.csv" , stringsAsFactors = FALSE, check.names = FALSE)


ggplot(ra, aes(`Date (C.E.)`, `Relative Abundance`)) +
  geom_bar(binwidth = 1.0, stat="identity") +
  coord_flip() +
  xlab("Date (C.E.)") +
  ylab("Relative Abundance") +
  ylim(0.0,0.2) +
  theme(
  legend.position="",
  axis.text.x = element_text(vjust=0.5, color="black", size=18),
  axis.text.y = element_text(vjust=0.5, color="black", size=18), 
  axis.title.y = element_text(vjust=1.0, color="black", size=18),
  axis.title.x = element_text(vjust=0.1, color="black", size=18),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(), 
  axis.line = element_line(colour = "black"),
  strip.text.x = element_text(size=15),
  strip.background = element_rect(colour="black", fill="white"))

fig_width <- 300 # play with this number
ggsave(filename = "Manuscript Figures/fig_6a.pdf",
       dpi = 600, units = "mm",
       height = fig_width/1.6, width =  fig_width)
  
```

```{r taxonomic_evenness}
# Calculate taxonomic evenness. Use the diversity function in vegan to 
# calculate the inverse of Simpson's per stratum. 

taxonomic_evenness <- ldply(ah[,-1],  function(i) diversity(i , index = "invsimpson"))
names(taxonomic_evenness) <- c("Date (C.E.)", "1/D")

# save as CSV 
write.csv(taxonomic_evenness, "data/fig6b.csv", row.names = FALSE)

# display table
kable(taxonomic_evenness)
```

```{r plot_1/D}
# Plot of 1/D. Use 1/D values that were previously 
# calculated to plot changing evenness through time.

even <- read.csv("data/fig6b.csv", stringsAsFactors = FALSE, check.names = FALSE)

 
ggplot(even, aes(`1/D`, `Date (C.E.)`, group=1)) +
  geom_point(size=7) +
  xlim(5.0,15.0) +
  xlab("1/D") +
  ylab("Date (C.E.)") +
  theme(
  legend.position="",
  axis.text.x = element_text(vjust=0.5, color="black", size=18),
  axis.text.y = element_text(vjust=0.5, color="black", size=18), 
  axis.title.y = element_text(vjust=1.0, color="black", size=18),
  axis.title.x = element_text(vjust=0.1, color="black", size=18),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), 
  panel.background = element_blank(), 
  axis.line = element_line(colour = "black"),
  strip.text.x = element_text(size=15),
  strip.background = element_rect(colour="black", fill="white"))

fig_width <- 300 # play with this number
ggsave(filename = "Manuscript Figures/fig_6b.png",
       dpi = 600, units = "mm",
       height = fig_width/1.6, width =  fig_width)
  
```

```{r plot_C_N_isotope}
#  Plot of carbon and nitrogen isotopic data from Arroyo Hondo Pueblo.

turkey <- read.csv("data/fig7.csv", stringsAsFactors = FALSE, check.names = FALSE)

ggplot(turkey, aes(x = d13C, y = d15N)) + 
  geom_point(aes(), size=3.5) + 
  xlab(expression(paste(delta^{13},'C'))) + 
  ylab(expression(paste(delta^{15},'N'))) + 
  xlim(-21.0,-5.0) +
  ylim(2.0,12.0) +
  theme(
  legend.position="",
  axis.text.x = element_text(vjust=0.5, color="black", size=18),
  axis.text.y = element_text(vjust=0.5, color="black", size=18), 
  axis.title.y = element_text(vjust=1.0, color="black", size=22),
  axis.title.x = element_text(vjust=0.1, color="black", size=22),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), 
  panel.background = element_blank(), 
  axis.line = element_line(colour = "black"))

# save plot
fig_width <- 300 # play with this number
ggsave(filename = "Manuscript Figures/fig_7.png",
       dpi = 600, units = "mm",
       height = fig_width/1.6, width =  fig_width)
```

```{r plot_C_C_isotope}
#  Plot of collagen and apatite carbon isotopic data from Arroyo Hondo Pueblo.

turkey2 <- read.csv("data/fig8.csv", stringsAsFactors = FALSE, check.names = FALSE)

ggplot(turkey2, aes(x = d13Cc, y = d13Ca)) + 
  geom_point(aes(), size=3.5) + 
  xlab(expression(paste(delta^{13},'C collagen'))) + 
  ylab(expression(paste(delta^{13},'C apatite'))) + 
  xlim(-18.0,-7.0) +
  ylim(-11,0) +
  theme(
  legend.position="",
  axis.text.x = element_text(vjust=0.5, color="black", size=18),
  axis.text.y = element_text(vjust=0.5, color="black", size=18), 
  axis.title.y = element_text(vjust=1.0, color="black", size=22),
  axis.title.x = element_text(vjust=0.1, color="black", size=22),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), 
  panel.background = element_blank(), 
  axis.line = element_line(colour = "black"))

# save plot
fig_width <- 300 # play with this number
ggsave(filename = "Manuscript Figures/fig_8.png",
       dpi = 600, units = "mm",
       height = fig_width/1.6, width =  fig_width)
```

```{r summary_isotope_data}
# Calculate means and standard deviations for carbon and 
# nitrogen isotopic data. 

summary_isotope_data <- apply(turkey[,-1], 2, function(x) c(mn=mean(x), sd=sd(x)))
kable(summary_isotope_data)

# Calculate means and standard deviations for collagen and 
# apatite carbon data. 

summary_isotope_data2 <- apply(turkey2[,-1], 2, function(x) c(mn=mean(x), sd=sd(x)))
kable(summary_isotope_data2)

```

```{r session_info}
# Computational environment and software dependencies for this analysis
sessionInfo()
```

LICENSE
 
The MIT License (MIT)

Copyright (c) 2015 Cyler Conrad & Ben Marwick

Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in
all copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
THE SOFTWARE.


